Capital Budget Funding in Toronto

### 2024-09-25 paper/paper.qmd
included table showing yearly and total funding for ward

@@

2.  Ward Profiles (2021): \\
This dataset provides comprehensive socioeconomic information for each ward based on the 2021 census. It was filtered to remain the variables which were most useful for this research. These are: Total Population, Median Household Income, Prevalence of Low Income Households, Educational Attainment, Immigration Status, Visible Minority Population. [may need to include others, depends on what I decide to go with in the end.]

@@

>>>>

```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace setup ####
library(dplyr)
library(tidyverse)
library(readr)
library(knitr)
library(kableExtra)
library(scales)

### Data handling ###
# Import cleaned data 
budget_data_clean <- read_csv("../data/analysis_data/2022-2031-budget_data.csv")

# Identify year columns
year_cols <- as.character(2022:2031)

# Reshape data to long format
budget_data_long <- budget_data_clean |>
  pivot_longer(
    cols = all_of(year_cols),
    names_to = "Year",
    values_to = "Funding"
  )

### Summing funding values ###

# Total overall funding per ward number 
total_funding_ward <- budget_data_clean |>
  group_by(`Ward Number`) |>
  summarise(
    Total_Funding = sum(across(all_of(year_cols)), na.rm = TRUE)
  ) |>
  arrange(desc(Total_Funding))

# Funding for each year per ward number 
funding_yearly_ward <- budget_data_long |>
  group_by(`Ward Number`, Year) |>
  summarise(Funding = sum(Funding, na.rm = TRUE)) |>
  arrange(`Ward Number`, Year)

# Total funding per program/agency per ward number 
total_funding_program_ward <- budget_data_clean |>
  group_by(`Ward Number`, `Program/Agency Name`) |>
  summarise(Total_Funding = sum(across(all_of(year_cols)), na.rm = TRUE)) |>
  arrange(`Ward Number`, desc(Total_Funding))

# Yearly funding per program/agency per ward number - maybe remove later
funding_yearly_program_ward <- budget_data_long |>
  group_by(`Ward Number`, `Program/Agency Name`, Year) |>
  summarise(
    Funding = sum(Funding, na.rm = TRUE)
  ) |>
  arrange(`Ward Number`, `Program/Agency Name`, Year)

```

@@

2.  Ward Profiles (2021): \\ This dataset provides comprehensive socioeconomic information for each ward based on the 2021 census. It was filtered to remain the variables which were most useful for this research. These are: Total Population, Median Household Income, Prevalence of Low Income Households, Educational Attainment, Immigration Status, Visible Minority Population. \[may need to include others, depends on what I decide to go with in the end.\]

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Yearly Funding per Ward Number with Total Column
funding_yearly_ward_table <- funding_yearly_ward |>
  pivot_wider(names_from = Year, values_from = Funding, values_fill = 0) |> 
  rowwise() |> 
  mutate(`Total Funding` = sum(c_across(all_of(year_cols)))) |> 
  ungroup() |> 
  mutate_at(vars(all_of(year_cols), `Total Funding`), ~ dollar(.)) |> 
  arrange(`Ward Number`)

funding_yearly_ward_table |> 
  kable(
    col.names = c("Ward Number", year_cols, "Total Funding"),
    caption = "Yearly Funding for Wards"
  ) |>

  kableExtra::kable_styling(bootstrap_options = "condensed", full_width = F, font_size = 6, position = "center")

```
